Peaks Over Threshold for Bursty Time Series

### 2018-04-16 article_springer/article_springer.Rmd
remove block maxima discussion

[^1]: (To see why, note that $X(\ell)$ is, in distribution, simply equal to 
  $J | J > \ell$, independent of any waiting time $W_k$.)

@@

Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top 
panel, where $W$ has a stable distribution with tail parameter 
$\beta =$ `r params$tail` (and skewness $1$ and location $0$), and 
where $J$ is from a standard Gumbel distribution. In the bottom panel, 
we plot a time series of solar flare intensities derived from a NASA
dataset [@HXRBS][^*]. Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear 
continuously. The threshold exceedances, however, appear to have the same 
statistical behaviour in both models. Observations below a threshold are 
commonly discarded in Extreme Value Theory; the MRP models 
interprets these observations as noise and filters them out.

@@

In classical extreme value theory, there are two main approaches for 
the inference on GEV distributions: Block
Maxima and Peaks-Over-Threshold (POT). Block 
Maxima subdivide the time line into 'blocks' of equal length, 
and the maximum of each block is taken as realization of the random 
variable $M_n$. 
This approach is not appropriate for CTRM data, because maxima in 
same-sized blocks will not be i.i.d.
As mentioned in Section 2, the POT approach *does* extract i.i.d. $X(\ell)$ 
(and $T(\ell)$), and can hence be directly used for the inference on 
the extreme values of $J_k$. 
Figure \ref{fig:GP-simulated} shows an example for the so-called 
"threshold-selection" plots for data drawn from the standard Gumbel 
distribution. 
For a range of thresholds, maximum likelihood estimators of the shape and 
scale parameters in the GP distribution are computed and plotted. 
Note that the scale parameter is transformed so that it appears independent
of the threshold [see e.g. Section 4.3.4 in @ColesBook].

```{r GPD-simulated-data, warning=FALSE, fig.cap="\\label{fig:GP-simulated}Stability plot for the shape and scale parameters of the generalized Pareto distribution. Here, the magnitudes are standard Gumbel, hence $\\xi = 0$ and $\\sigma = 1$. Plots were generated using the POT package in R [@R-POT]."}
par(mfrow = c(1,2))
JJ <- coredata.ctrm(sim_ctrm)
u.range <- quantile(x = JJ, probs = c(0.75, 0.95))
POT::tcplot(JJ, u.range = u.range)
# extRemes::threshrange.plot(mrp("getMags"), type = "GP", set.panels = FALSE)
```

For the remainder of this article, we focus on statistical inference 
for the inter-exceedance times $T(\ell)$ of the threshold crossings. 
In our setting, the inter-exceedance times follow a fractional 
Poisson process instead of a standard Poisson process, which is how our
approach generalizes the standard POT approach.

>>>>

[^1]: To see why, note that $X(\ell)$ is, in distribution, simply equal to 
  $J | J > \ell$, independent of any waiting time $W_k$.

@@

Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top 
panel, where $W$ has a stable distribution with tail parameter 
$\beta =$ `r params$tail` (and skewness $1$ and location $0$), and 
where $J$ is from a standard Gumbel distribution. In the bottom panel, 
we plot a time series of solar flare intensities derived from a NASA
dataset [@HXRBS][^*]. Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear 
continuously. The threshold exceedances, however, appear to have the same 
statistical behaviour in both models. Observations below a threshold are 
commonly discarded in Extreme Value Theory; likewise, our MRP model
interprets these observations as noise and filters them out.

@@
